---
title: "Covid Data by State"
author: Victoria York
date: Sept 26 2021
output: html_notebook
---
The goal of this program is to show the number of deaths from covid per day in the states of New York, Florida, Connecticut, Massachusetts, New Jersey, Texas, California, Washington, Illinois and Idaho. 
We took data from the New York Times and downloaded the file into this program. In the first code chunck we will be assinging this code the varible name of covid. 
```{r}
library(ggplot2)
library(readr)
library(dplyr)
library(glue)
library(lubridate)
library(scales)
covid <- read_csv("covid_data_cdc.csv")
```

Next we will create a varible that will hold a vector of the names of the states that we wish to include in our data. 

```{r select-states}
states <- c(
  "NY",
  "FL",
  "CT",
  "MA",
  "NJ",
  "TX",
  "CA",
  "WA",
  "IL",
  "ID"
)
```

In the following code chunk we filter the data frame for deaths in the states we specified above and calculate cumulative number of deaths.

```{r}
state_data <- covid %>% 
  # only includes states within our state vector
  filter(state %in% states) %>%
  # only includes positive new death values
  filter(new_death >= 0) %>%
  group_by(state, submission_date) %>%
  # arrange by date in ascending order
  arrange(submission_date) %>%
  ungroup()

state_data$submission_date <- as.Date(state_data$submission_date, "%m/%d/%Y")
```

The following visualisation shows the number of deaths vs. days elapsed in various states in the US.

```{r}
ggplot(data = state_data,
       mapping = aes(x = submission_date, 
                     y = new_death, 
                     color = state,
                     group = state,
                     label = state)) +
  # represent cumulative cases with lines
  geom_line(size = 0.7, alpha = 0.8) +
  # use pretty colors
  scale_color_viridis_d() +
  # better formatting for y-axis
  scale_y_continuous(labels = label_comma()) +
  # use minimal theme
  theme_minimal() +
  # customize labels
  labs(
    x = "Date",
    y = "Number of deaths",
    title = "Deaths from COVID-19, Selected States",
    caption = "Source: https://data.cdc.gov/Case-Surveillance/United-States-COVID-19-Cases-and-Deaths-by-State-o/9mfq-cb36/data"
  )
```

To further better this data I think it would be useful to make the graphs graph the percentage of covid deaths in the population. This would show porortionaly which states are struggling the most. 
