---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Victoria York"
date: "Oct 8 2021"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
library(readr)
```

```{r load-data, message = FALSE}
states <- read_csv("states.csv")
dennys <- dsbox::dennys
laquinta <- dsbox::laquinta
```

### Exercise 1

```{r}
nrow(dennys)
ncol(dennys)
```
dimentions: rows: 1643, columns: 6. 
each row represents a Denny's location. 
the varibles are adress, city, state, zip, longitude, latitude

### Exercise 2

```{r}
nrow(laquinta)
ncol(laquinta)
```
dimention: rows: 909, columns: 6
each row represents a La Quinta's location. 
the varibles are adress, city, state, zip, longatude, latitude 

### Exercise 3

La Quinta has locations in: Canada, Mexico, New Zealand, Honduras, Turkey, United Arab Emirates, Chile and Colombia. But Denny's does not have any locations outside of the US. 

### Exercise 4

Have the location split up into country and province. Then we could filter by the country not being the US. 

### Exercise 5

```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
No there are no Denny's outside the US 

### Exercise 6

```{r}
dennys <- dennys %>%
  mutate(country = "United States")
```

# Exercise 7

Canada: 
British Columbia - BC
Ontario - ON

Mexico:
AG
VE
PU
SL

New Zealand:
Not in data set

Honduras:
FM

Turkey:
None were in data set

United Arab Emirates:
Not in data set

Chile:
Not in dats set

Colombia:
ANT

# Exercise 8

```{r}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("AG", "VE", "PU", "SL") ~ "Mexico",
    state == "FM" ~ "Honduras"
  ))

laquinta <- laquinta %>%
  filter(country == "United States")
```

# Exercise 9

```{r}
numofdennys <- dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

```


#Exercise 10

```{r}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dennys, laquinta)

ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```

#Exercise 11

```{r}
northcar <- dn_lq %>% filter(state == "NC")

ggplot(northcar, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```

#Exercise 12

```{r}
texas <- dn_lq %>% filter(state == "TX")

ggplot(texas, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```

Note: I had time to do the whole lab except ran out of time to play with the alpha function. I did try to get it to work but was unsucessful. 



